### Additional features

The demo above covers the basics: acting on the page, moving the camera, annotating, and captioning. paparazzi can also:

- **Draw attention to what matters.** Callouts with arrows (`pz_annotate_callout()`) and spotlights that dim everything else (`pz_annotate_spotlight()`) lead your viewer's eye to the part of the page you're talking about. In a recording, paparazzi animates annotations in and keeps them on their elements as the page scrolls or changes.
- **Set defaults once.** The staging functions set defaults for the whole page, so you don't repeat the same arguments in every step. `pz_stage()` sets the cursor's speed and size, the typing speed, and the pause after each step, `pz_stage_annotate()` sets how annotations look, and `pz_stage_frame()` sets the default framing.
- **Control what's in the shot.** With `pz_frame()`, you choose exactly what a video or screenshot shows: crop to one or more elements, add padding, and set an aspect ratio, like 16:9 for slides or 9:16 for a vertical video.
- **Show keyboard shortcuts.** When you press keys with `pz_act_press()`, paparazzi can show them in the recording as on-screen keycaps.
- **Write subtitles.** For MP4 and WebM videos, paparazzi can write your captions to a WebVTT subtitle file, instead of or as well as burning them into the video.
- **Use your own fonts.** `pz_stage_fonts()` loads fonts from Google Fonts, Bunny Fonts, or a file for captions, callouts, and badges.
- **Hide private data.** Cover sensitive text with `pz_annotate_redact()` before you start recording, and it stays hidden from the very first frame.
- **Set up the browser the way you want it.** `pz_device()` gives you one place to configure the browser: screen size, phone emulation, dark mode, zoom, reduced motion, locale, and time zone.

### Screenshots

You can use the same annotations, framing, and devices in screenshots. Here's the task tracker on a phone in dark mode, with a callout:

```{r should-update-phone}
#| include: false

# Same staleness gate as the hero: only recapture when the image is
# missing or the source fragment has changed since.
should_update_phone <- identical(Sys.getenv("IN_INDEX", ""), "true") && {
  img <- knitr::fig_chunk("phone", "png", 1)
  !file.exists(img) || file.mtime(file.path(root(), "man/fragments/additional-features.Rmd")) > file.mtime(img)
}
```

```{r phone, eval = should_update_phone}
#| include: false
#| out.width: 300px

phone <- pz_open(
  pz_example("tasks"),
  width = 390,
  height = 700,
  mobile = TRUE,
  color_scheme = "dark"
)

phone |>
  pz_annotate_callout(
    "Tap Done to finish a task",
    target = pz_loc(".task-done", which = "first"),
    side = "bottom"
  ) |>
  pz_screenshot()
```

```{r, eval = should_update_phone}
#| include: false

pz_close(phone)
```

<table class="table-borderless">
<tr>
<td valign="top" style="width: 100%; max-width: 0;">

```{r}
#| label: phone-code
#| eval: false
#| ref.label: phone
```

</td>
<td valign="top" width="300" style="min-width: 260px;">

```{r}
#| echo: false
#| out.width: 300px
#| fig.alt: "The task tracker on a narrow phone screen in dark mode. A red-bordered callout under the first task's Done button reads Tap Done to finish a task."

path <- if (identical(Sys.getenv("IN_INDEX", ""), "true")) {
  "images/readme-phone-1.png"
} else {
  "https://raw.githubusercontent.com/posit-dev/paparazzi/main/pkgdown/assets/images/readme-phone-1.png"
}

knitr::include_graphics(path, error = FALSE)
```

</td>
</tr>
</table>
