## Features

### Point the viewer at what matters

Zoom and pan the camera to a form, a button, or a whole panel with `pz_camera()`. While it's zoomed in, the camera pans to keep each click and keystroke in view, so nothing happens out of the shot. The camera only exists in the video: it never scrolls or zooms the page itself. Outline, circle, underline, or highlight elements with `pz_annotate()`, with numbered badges if you want them. Add callouts with arrows (`pz_annotate_callout()`), or dim everything except one element with a spotlight (`pz_annotate_spotlight()`). In a recording, annotations draw, pop, fade, slide, or wipe in. They follow their elements as the page scrolls or changes.

### Make it look like a person is using the page

The cursor glides between elements, changes shape over text fields and links, and presses down when it clicks. Typing arrives one character at a time at a natural pace. `pz_stage()` sets the cursor's speed and size, the typing speed, and the pause after each step.

### Tell the story

Captions sit at the top or bottom of the video and change as you go. You can burn them into the video or, for MP4 and WebM, write them to a WebVTT subtitle file alongside it. Keyboard shortcuts pressed with `pz_press()` can appear on screen as keystrokes. `pz_record_hold()` gives viewers time to read, and `pz_record_pause()` cuts out anything they don't need to see.

### Frame it for wherever it's going

Crop a video or a screenshot to one element, or to several, with padding, using `pz_frame()`. Set an aspect ratio, like 16:9 for slides or 9:16 for a vertical video. Save MP4, WebM, or GIF files, and scale them down to a size that suits a README.

### Make every version from one script

Record or capture the same steps on a phone, in dark mode, zoomed in, or in another locale or time zone. `pz_device()` switches between them. Cover private data with `pz_annotate_redact()` before recording starts, so it stays hidden from the very first frame.

Screenshots use the same annotations and framing. Here's the task tracker on a phone in dark mode, with a callout:

```{r phone, eval = identical(Sys.getenv("IN_INDEX", ""), "true")}
#| include: false
#| out.width: 300px

phone <- pz_open(
  pz_example("tasks"),
  width = 390,
  height = 700,
  mobile = TRUE,
  color_scheme = "dark"
)

phone |>
  pz_annotate_callout(
    "Tap Done to finish a task",
    target = pz_loc(".task-done", which = "first"),
    side = "bottom"
  ) |>
  pz_screenshot()
```

```{r, eval = identical(Sys.getenv("IN_INDEX", ""), "true")}
#| include: false

pz_close(phone)
```

```{r}
#| echo: false
#| results: asis

path <- if (identical(Sys.getenv("IN_INDEX", ""), "true")) {
  "images/readme-phone-1.png"
} else {
  "https://raw.githubusercontent.com/posit-dev/paparazzi/main/pkgdown/assets/images/readme-phone-1.png"
}

cat(sprintf(
  '<img src="%s" alt="%s" />',
  path,
  "The task tracker on a narrow phone screen in dark mode. A red-bordered callout under the first task's Done button reads Tap Done to finish a task."
))
```

Like the recording, a screenshot at the end of a knitted chunk becomes a figure in the document.
