---
output: github_document
---

<!-- index.md is generated from index.Rmd. Please edit that file -->

```{r}
#| include: false

root <- function(dir = getwd()) {
  if (file.exists(file.path(dir, "DESCRIPTION"))) {
    return(dir)
  } else if (dirname(dir) == dir) {
    return(NULL)
  }
  root(dirname(dir))
}

withr::local_envvar(IN_INDEX = "true")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = file.path(root(), "pkgdown/assets/images/readme-")
)
```

```{r}
#| child: ../man/fragments/headline.Rmd
```

```{r}
#| child: ../man/fragments/hero.Rmd
```

```{r}
#| child: ../man/fragments/about.Rmd
```

```{r}
#| child: ../man/fragments/installation.Rmd
```

```{r}
#| child: ../man/fragments/hero-about.Rmd
```

```{r}
#| child: ../man/fragments/features.Rmd
```

```{r}
#| child: ../man/fragments/shiny.Rmd
```

```{r}
#| child: ../man/fragments/testing.Rmd
```

## Learn more

- [Get started](articles/paparazzi.html) builds a demo video one step at a time, from opening a page to framing the finished recording.
- [Make a three-part app walkthrough](articles/walkthrough.html) records a series of short videos with camera moves, callouts, captions, and redactions.
- [Annotate pages](articles/annotations.html) tours every kind of mark, callout, spotlight, redaction, and caption.
- [Shiny apps](articles/shiny.html) covers starting apps, waiting for the server, setting inputs, and testing.
- The [function reference](reference/) lists every function by task.
