---
title: "Annotate pages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Annotate pages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Annotations draw attention to parts of a page: an outline around a field, a callout explaining a button, a cover over something private. They sit on an overlay above the page and follow their elements as the page scrolls or changes layout, and they show up in screenshots and recordings until you clear them. They belong to the current document, so navigating away removes them.

This article tours every kind of annotation on the task tracker bundled with paparazzi. The [Get started article](paparazzi.html) covers selectors and scopes; the [walkthrough](walkthrough.html) uses annotations in longer recordings. You'll need a Chromium-based browser. The GIF recordings at the end need gifski, and the keystroke callouts in a GIF also need av.

```{r setup}
library(paparazzi)
page <- pz_open(
  pz_example("tasks"),
  width = 720,
  height = 800,
  color_scheme = "light"
)
```

## Marks

[`pz_annotate()`](../reference/pz_annotate.html) draws a mark on every element matching `target`, which is a selector, a [`pz_loc()`](../reference/pz_loc.html) spec, or a list of targets. With no target, it marks the current scope. There are four mark types:

- `"box"` (the default) outlines the element.
- `"circle"` draws an ellipse around it, with room for things that aren't rectangular.
- `"underline"` strokes along the element's bottom edge.
- `"highlight"` lays a translucent fill over the element, blended with the page.

```{r mark-types, fig.alt = "The task tracker with four labeled marks: a box around the title field, a circle around the Add button, an underline under the first task, and a highlight over the second task."}
page |>
  pz_annotate("#task-title", type = "box", label = "box") |>
  pz_annotate("#add-task", type = "circle", label = "circle") |>
  pz_annotate(
    pz_loc(".task-title", which = "first"),
    type = "underline",
    label = "underline"
  ) |>
  pz_annotate(
    pz_loc(".task", which = 2),
    type = "highlight",
    label = "highlight"
  )
page |> pz_screenshot(frame = pz_frame("main", pad = 16))
```

Here and below, each screenshot passes a [`pz_frame()`](../reference/pz_frame.html) so the still is cropped to what it discusses. Marks stay on the page after they're drawn, so clear them between examples:

```{r}
page |> pz_annotate_clear()
```

### Badges

`label` adds a small badge to each mark. `label = TRUE` numbers the matches in document order, which is handy when a screenshot refers to "item 3":

```{r labels, fig.alt = "The three filter links, All, Open, and Done, boxed and numbered 1, 2, and 3."}
page |> pz_annotate(".filters a", label = TRUE)
page |>
  pz_screenshot(
    frame = pz_frame(".filters", pad = 16, target_box = "annotated")
  )
```

A single string or number repeats on every match instead, so `label = "A"` or `label = 1` puts the same badge everywhere the target matches.

```{r}
page |> pz_annotate_clear()
```

### Padding, color, and fonts

`pad` grows the mark beyond the element's box: one number for all sides, or `c(top, right, bottom, left)`. `color` accepts any CSS color for the outline and badge, and `font_family` and `font_size` style the badge text. All three default to the page's staged settings, which [Styling defaults] covers below.

```{r pad-color, fig.alt = "The new-task form with a loose amber highlight around the title field and a blue box with a monospace A badge around the priority dropdown."}
page |>
  pz_annotate("#task-title", type = "highlight", pad = 6, color = "#f59e0b") |>
  pz_annotate(
    "#task-priority",
    label = "A",
    pad = c(2, 6, 2, 6),
    color = "#2563eb",
    font_family = "monospace",
    font_size = 16
  )
page |> pz_screenshot(frame = pz_frame("#new-task", pad = 24))
page |> pz_annotate_clear()
```

## Callouts

[`pz_annotate_callout()`](../reference/pz_annotate_callout.html) draws a dark bubble with white text next to each matching element, repeating the same `text` for every match. `side` picks where the bubble goes: a cardinal side (`"top"`, `"right"`, `"bottom"`, `"left"`) or a diagonal like `"top right"`. The default, `NULL`, chooses the cardinal side with the most room. A bubble is kept inside the viewport even if that moves it off the requested side.

`arrow = FALSE` drops the arrow for a tooltip-style bubble, and `label` adds a badge just like on marks.

```{r callouts, fig.alt = "The new-task form with a numbered callout above the title field pointing down at it, and a tooltip-style callout without an arrow below the priority dropdown."}
page |>
  pz_annotate_callout(
    "Name the task",
    target = "#task-title",
    side = "top",
    label = 1
  ) |>
  pz_annotate_callout(
    "Priority is optional",
    target = "#task-priority",
    side = "bottom",
    arrow = FALSE
  )
page |>
  pz_screenshot(
    frame = pz_frame("#new-task", pad = 12, target_box = "annotated")
  )
page |> pz_annotate_clear()
```

The frame's `target_box = "annotated"` keeps the crop wide enough to include the bubbles; [Framing stills] explains what that does.

## Spotlight

[`pz_annotate_spotlight()`](../reference/pz_annotate_spotlight.html) dims the whole page except for a rounded cutout around each target, so the reader's eye goes straight to what matters. `dim` sets the opacity of the black scrim from 0 (transparent) to 1 (black), defaulting to 0.6, and `pad` grows the cutouts. A list of targets gets one cutout each.

```{r spotlight, fig.alt = "The task tracker dimmed to near-black, with bright cutouts around the title field and the Add button."}
page |>
  pz_annotate_spotlight(
    list("#task-title", "#add-task"),
    pad = 8,
    dim = 0.75
  )
page |> pz_screenshot(frame = pz_frame("main", pad = 24))
```

There is only one spotlight per page: drawing a new one replaces the old. Clear it by its reserved id or by clearing everything:

```{r}
page |> pz_annotate_clear(id = "spotlight")
```

## Redaction

[`pz_annotate_redact()`](../reference/pz_annotate_redact.html) covers elements so their contents don't reach the screenshot or recording. `method = "fill"` (the default) paints an opaque cover, near-black unless you pass a `color`:

```{r redact-fill, fig.alt = "The high-priority task titles covered by solid near-black bars."}
page |> pz_annotate_redact('[data-priority="high"] .task-title', method = "fill")
page |>
  pz_screenshot(
    frame = pz_frame('[data-priority="high"] .task-title', pad = 24, target_box = "annotated")
  )
page |> pz_annotate_clear()
```

`method = "blur"` blurs what's underneath instead:

```{r redact-blur, fig.alt = "The high-priority task titles blurred so the text is unreadable but its shape is visible."}
page |> pz_annotate_redact('[data-priority="high"] .task-title', method = "blur")
page |>
  pz_screenshot(
    frame = pz_frame('[data-priority="high"] .task-title', pad = 24, target_box = "annotated")
  )
page |> pz_annotate_clear()
```

Use fill for real secrets: blur can leave short text partly legible, and a fill stays opaque even if you give it a translucent color. A redaction covers each element's border box plus `pad`, but not content overflowing the element, so redact the element that actually overflows or add padding. Redactions follow axis-aligned overflow clipping (not custom `overflow-clip-margin`) and hide if their target disconnects or stops rendering. Redactions draw instantly, so add them *before* starting a recording: frames captured earlier can already contain the text.

## Captions

[`pz_annotate_caption()`](../reference/pz_annotate_caption.html) adds a caption in screen space: a dark panel floating at the `side = "bottom"` (the default) or `"top"` edge of the capture. It's not attached to an element. It persists through navigation and across recordings, and it's composited onto stills. Newlines are preserved.

```{r caption, fig.alt = "The task tracker with a dark caption panel centered along the bottom edge, reading: Seven tasks, two clicks from done."}
page |> pz_annotate_caption("Seven tasks, two clicks from done")
page |> pz_screenshot()
```

The caption sits at the edge of the *viewport*, so a full-page still shows it as above. In a still cropped to a frame, the caption appears wherever it happened to be on screen, which is rarely what you want; captions belong to full-page stills and recordings.

Calling `pz_annotate_caption()` again replaces the caption. Remove it by its reserved id:

```{r}
page |> pz_annotate_clear(id = "caption")
```

## Clearing and replacing

[`pz_annotate_clear()`](../reference/pz_annotate_clear.html) removes one annotation by `id`, or every annotation when `id` is `NULL`. The reserved ids `"spotlight"` and `"caption"` clear those types; your own marks, callouts, and redactions take any nonempty id. Reusing an id replaces that annotation instead of stacking a second one:

```{r replace, fig.alt = "The new-task form with only a circle around the Add button; the box that was around the title field is gone."}
page |> pz_annotate("#task-title", id = "mark")
page |> pz_annotate("#add-task", type = "circle", id = "mark")
page |> pz_screenshot(frame = pz_frame("#new-task", pad = 16))
page |> pz_annotate_clear()
```

Only the circle remains: the second `id = "mark"` call replaced the box. During an active recording, clearing plays the annotation's reveal in reverse; in stills, annotations disappear instantly.

## Styling defaults

Every annotation defaults to the page's staged color, font family, and badge size. Change them for all later annotations with [`pz_stage_annotate()`](../reference/pz_stage_annotate.html), and pass `NULL` to restore a default (crimson `#e11d48`, sans-serif, 14 CSS pixels):

```{r stage-style, fig.alt = "The Add button boxed in blue with a large monospace A badge."}
page |>
  pz_stage_annotate(
    color = "#2563eb",
    font_family = "monospace",
    font_size = 18
  ) |>
  pz_annotate("#add-task", label = "A")
page |>
  pz_screenshot(
    frame = pz_frame("#new-task", pad = 8, target_box = "annotated")
  )
```

```{r}
page |> pz_annotate_clear()
page |>
  pz_stage_annotate(
    color = NULL,
    font_family = NULL,
    font_size = NULL
  )
```

Per-call `color`, `font_family`, and `font_size` override the staged values for that call.

## Framing stills

A screenshot's frame is measured from the target's elements. Badges and callouts paint *outside* those elements, so the default `target_box = "element"` can crop them off. `target_box = "annotated"` measures the elements plus their attached marks, callouts, and spotlight cutouts. Here's the difference on the same callout:

```{r frame-setup}
page |> pz_annotate_callout("The list so far", target = "h1", side = "right")
```

```{r frame-element, fig.alt = "A tight crop of the Tasks heading alone; the callout next to it is cut off."}
page |> pz_screenshot(frame = pz_frame("h1", pad = 16))
```

```{r frame-annotated, fig.alt = "A crop of the Tasks heading that includes the full callout bubble to its right."}
page |>
  pz_screenshot(
    frame = pz_frame("h1", pad = 16, target_box = "annotated")
  )
```

```{r}
page |> pz_annotate_clear()
```

Redactions and annotations on unrelated elements are never included. `target_box = "annotated"` applies to stills; a recording's home frame always uses the element box.

## Reveals

So far every mark has appeared fully drawn, because reveal animations only play during an active, unpaused recording. `reveal` chooses the animation: `"fade"`, `"draw"` (the stroke draws itself), `"pop"`, `"slide"`, `"wipe"` (sweeps clockwise from 12 o'clock), or `"none"`. The defaults are fade for boxes, draw for the other mark types, pop for callouts, and fade for spotlights. Clearing an annotation plays its reveal in reverse. This recording shows a box fading in, a circle drawing itself, and a callout popping in, then all three reversing out:

```{r reveals}
page |>
  pz_record_start(
    frame = pz_frame("#new-task", pad = 40, when = "start"),
    scale = 0.6,
    hold = c(0.3, 0.5)
  ) |>
  pz_annotate("#task-title", type = "box") |>
  pz_record_hold(0.4) |>
  pz_annotate("#task-priority", type = "circle") |>
  pz_record_hold(0.4) |>
  pz_annotate_callout(
    "Then press Add",
    target = "#add-task",
    side = "bottom"
  ) |>
  pz_record_hold(1) |>
  pz_annotate_clear()
```

The recording stops when the chain ends, and with no path it's written into the article as an MP4. Pass a path to choose the file, or `format = "gif"` for a GIF.

## Keystroke callouts

One more callout exists only in recordings: [`pz_act_press(show_keys = ...)`](../reference/pz_act_press.html) shows the keys being pressed as keycaps. `"words"` spells out modifier names, `"mac"` uses Mac symbols, and `"both"` shows `Mod` as `Ctrl / ⌘`. The default is `"none"`; set [`pz_stage(show_keys = ...)`](../reference/pz_stage.html) to change it for a whole page.

```{r show-keys}
page |>
  pz_record_start(
    frame = pz_frame("#new-task", pad = 24, when = "start"),
    scale = 0.6
  ) |>
  pz_act_click("#task-title") |>
  pz_act_type("Water the dog", target = "#task-title") |>
  pz_act_press("Enter", show_keys = "mac") |>
  pz_expect_text(
    "Water the dog",
    target = pz_loc(".task-title", which = "first")
  )
```

Keystroke callouts appear only in recordings, never in stills, and each `pz_act_press()` call replaces the previous callout.

Finally, close the page to shut down the browser:

```{r}
pz_close(page)
```

For annotations in a longer narrative recording, see the [three-part app walkthrough](walkthrough.html); the [function reference](../reference/index.html#annotations) lists every annotation function with its arguments.
