---
title: "Annotate pages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Annotate pages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

In this article, we'll draw every kind of annotation paparazzi offers on [the task tracker bundled with paparazzi](../reference/pz_example.html): marks, callouts, spotlights, redactions, and captions. Along the way, we'll style them, fit them into framed stills, and animate them in a recording, so you can choose the right annotation for each screenshot or video.

Annotations draw attention to parts of a page: an outline around a field, a callout explaining a button, a cover over something private. paparazzi draws most annotations into the page itself, attached to an element. That's how they follow their elements as the page scrolls or changes layout, and why they show up in screenshots as well as recordings.

Annotations stay on the page until you clear them, and they belong to the current document, so paparazzi removes them when the page navigates away. Captions are the exception: paparazzi adds them to each screenshot and recording instead of drawing them into the page, so they stay through navigation, as [Captions] explains.

If you're new to paparazzi, read the [Get started article](paparazzi.html) first: it introduces selectors, location specs, and scopes, which annotations use to pick their elements. Afterward, the [in-depth app walkthrough](demo-video.html) combines annotations in a longer recording. You'll need a Chromium-based browser, and the recordings at the end need the av package to encode MP4 files.

`pz_open()` starts a headless browser and opens the page, and `pz_example()` returns the path to the task tracker that comes with paparazzi:

```{r setup}
library(paparazzi)

page <- pz_open(
  pz_example("tasks"),
  width = 720,
  height = 800,
  color_scheme = "light"
)
```

## Marks

`pz_annotate()` takes a `target` selector, [location spec](../reference/pz_loc.html), or list of targets and draws an annotation mark on every matching element. With no target, it marks the current scope -- the element that `pz_find()` narrowed the chain to -- or the page body.

You choose one of four marks with the `type` argument:

- `"box"` (the default) outlines the element.
- `"circle"` draws an ellipse around it, with room for elements that aren't rectangular.
- `"underline"` strokes along the element's bottom edge.
- `"highlight"` lays a translucent fill over the element, blended with the page.

Let's draw one of each, with a `label` naming its type:

```{r mark-types, fig.alt = "The task tracker with four labeled marks: a box around the title field, a circle around the Add button, an underline under the File tax return task, and a highlight over the Water the plants task."}
page |>
  pz_annotate("#task-title", type = "box", label = "box") |>
  pz_annotate("#add-task", type = "circle", label = "circle") |>
  pz_annotate(
    pz_loc(".task", has_text = "File tax return"),
    type = "underline",
    label = "underline"
  ) |>
  pz_annotate(
    pz_loc(".task", has_text = "Water the plants"),
    type = "highlight",
    label = "highlight"
  )
page |> pz_screenshot(frame = pz_frame("main", pad = 16))
```

Each screenshot in this article passes a `pz_frame()`, so paparazzi crops the still to the part of the page we're discussing. Marks stay on the page after paparazzi draws them, so we'll clear them before each new example:

```{r}
page |> pz_annotate_clear()
```

paparazzi only draws a mark where its target can be seen. When a task row scrolls out of view inside the scrolling task list, paparazzi cuts its mark off at the list's edge instead of drawing it over whatever sits below the list. On the sides where the target isn't cut off, the outline keeps its padding, even past the list's edge. paparazzi hides the mark, and its badge, when the target is scrolled entirely out of the list, hidden, or removed from the page.

### Label badges

`label` adds a small badge to each mark. `label = TRUE` numbers the matches in document order, which helps when your text refers to "item 3":

```{r labels, fig.alt = "The three filter links, All, Open, and Done, boxed and numbered 1, 2, and 3."}
page |> pz_annotate(".filters a", label = TRUE)
page |>
  pz_screenshot(
    frame = pz_frame(".filters", pad = 16, target_box = "annotated")
  )
```

The badges sit outside the filter links, so this frame uses `target_box = "annotated"` to fit the links together with their marks and badges. [Framing stills] shows the difference.

A single string or number repeats on every match instead, so `label = "A"` or `label = 1` puts the same badge on each element the target matches.

```{r, include = FALSE}
page |> pz_annotate_clear()
```

### Padding, color, and fonts

`pad` grows the mark beyond the element's box, with one number for all sides or `c(top, right, bottom, left)`. `color` sets the outline and badge to any CSS color, and `stroke_width` sets the outline's thickness in CSS pixels. `font_family` and `font_size` style the badge text.

paparazzi fills each badge with the mark's color and writes its text in white; `label_fill` and `label_text_color` change that for one call. When you leave out the color, stroke width, or fonts, paparazzi uses the page's defaults, which you can change with `pz_stage_annotate()`, as [Styling defaults] explains.

Here, an amber highlight covers the title field, and a blue box with a monospace badge fits snugly around the Add button:

```{r pad-color, fig.alt = "The new-task form with a loose amber highlight around the title field and a blue box around the Add button, with a monospace badge reading #add-task."}
page |>
  pz_annotate("#task-title", type = "highlight", pad = 6, color = "#f59e0b") |>
  pz_annotate(
    "#add-task",
    label = "#add-task",
    pad = c(2, 6, 2, 6),
    color = "#2563eb",
    font_family = "monospace",
    font_size = 16
  ) |>
  pz_screenshot(
    frame = pz_frame("#new-task", pad = 24, target_box = "annotated")
  )
```

```{r, include = FALSE}
page |> pz_annotate_clear()
```

## Callouts

`pz_annotate_callout()` draws a bubble of text next to each matching element, repeating the same `text` for every match. `side` picks where the bubble goes: a side (`"top"`, `"right"`, `"bottom"`, `"left"`) or a corner like `"top right"`. When you leave `side` out, paparazzi picks the side with the most room.

paparazzi keeps every bubble inside the viewport, even if that moves it off the side you asked for, and it never cuts a bubble off. A leader line runs from the bubble to the element, with an arrow at the element's end. `leader = FALSE` drops the line for a tooltip-style bubble, and `label` adds a badge, just as it does for marks.

```{r callouts, fig.alt = "The new-task form with a numbered callout above the title field pointing down at it, and a tooltip-style callout without an arrow below the priority dropdown."}
page |>
  pz_annotate_callout(
    "Name the task",
    target = "#task-title",
    side = "top",
    label = 1
  ) |>
  pz_annotate_callout(
    "Priority is optional",
    target = "#task-priority",
    side = "bottom",
    leader = FALSE
  ) |>
  pz_screenshot(
    frame = pz_frame("#new-task", pad = 12, target_box = "annotated")
  )
```

```{r, include = FALSE}
page |> pz_annotate_clear()
```

Like a mark, a callout hides while its target is hidden or scrolled out of view inside a scrolling container. When only part of the target is in view, the leader points at that part.

### Leaders and bubble colors

`leader` also takes a decoration for each end of the line. `start` is the bubble's end and `end` is the element's end, and each can be `"none"`, `"arrow"`, `"dot"`, or `"bar"`. An end you leave out gets `"none"`. `distance` sets the gap between the bubble and the element in CSS pixels: 24 by default, or 8 for a bubble without a line. `stroke_width` sets the line's thickness, and paparazzi sizes the decorations to match.

paparazzi draws dark bubbles with white text unless you choose other colors. `fill` and `text_color` change the bubble's background and text, while `color` still sets its border and the line. Here, a long line with a dot at the bubble leads from a pale yellow bubble to the Add button:

```{r callout-styles, fig.alt = "The new-task form with a pale yellow callout below the Add button, joined to it by a long line with a dot at the bubble and an arrow at the button."}
page |>
  pz_annotate_callout(
    "Add it to the list",
    target = "#add-task",
    side = "bottom",
    leader = c(start = "dot", end = "arrow"),
    distance = 40,
    fill = "#fef3c7",
    text_color = "#171717"
  ) |>
  pz_screenshot(
    frame = pz_frame("#new-task", pad = 12, target_box = "annotated")
  )
```

```{r, include = FALSE}
page |> pz_annotate_clear()
```

## Spotlight

`pz_annotate_spotlight()` dims the whole page except for a rounded cutout around each target, so the viewer's eye goes straight to what matters. `dim` sets how dark the page gets, from 0 (no dimming) to 1 (black), and defaults to 0.6. `pad` grows the cutouts, and a list of targets gets one cutout each:

```{r spotlight, fig.alt = "The task tracker dimmed to near-black, with bright cutouts around the title field and the Add button."}
page |>
  pz_annotate_spotlight(
    list("#task-title", "#add-task"),
    pad = 8,
    dim = 0.75
  ) |>
  pz_screenshot(frame = pz_frame("main", pad = 24))
```

paparazzi only opens a cutout over the part of a target that's in view, so a target scrolled out of a scrolling container gets no cutout.

A page has one spotlight at a time: drawing a new spotlight replaces the old one. paparazzi gives the spotlight the reserved id `"spotlight"`, so you can clear it alone without clearing other annotations:

```{r}
page |> pz_annotate_clear(id = "spotlight")
```

## Redaction

`pz_annotate_redact()` covers elements so their contents stay out of the screenshot or recording. We'll treat the titles of the high-priority tasks as private. With `method = "fill"`, the default, paparazzi paints an opaque cover, near-black unless you pass a `color`:

```{r redact-fill, fig.alt = "The high-priority task titles covered by solid near-black bars."}
private <- '[data-priority="high"] .task-title'

page |>
  pz_annotate_redact(private, method = "fill") |>
  pz_screenshot(
    frame = pz_frame(private, pad = 24, target_box = "annotated")
  )
```

```{r, include = FALSE}
page |> pz_annotate_clear()
```

With `method = "blur"`, paparazzi blurs what's underneath instead:

```{r redact-blur, fig.alt = "The high-priority task titles blurred so the text is unreadable but its shape is visible."}
page |>
  pz_annotate_redact(private, method = "blur") |>
  pz_screenshot(
    frame = pz_frame(private, pad = 24, target_box = "annotated")
  )
```

```{r, include = FALSE}
page |> pz_annotate_clear()
```

Use fill for real secrets. Blur can leave short text partly legible, while a fill stays opaque even if you give it a translucent color.

A redaction covers each element's border box plus `pad`, but not content that overflows the element, so redact the element that actually overflows or add padding. Like marks, redactions stop at the edges of scrolling containers and hide while their target is hidden or removed.

paparazzi draws redactions instantly, even during a recording, but a frame captured before the redaction can already show the text. Add redactions *before* you start a recording, or while it's paused, as the [in-depth app walkthrough](demo-video.html#hide-private-details) does.

## Captions

`pz_annotate_caption()` adds a caption to the screen instead of to an element: a dark panel along the bottom edge of the capture, or the top edge with `side = "top"`. paparazzi adds the caption to each still and recording itself instead of drawing it into the page, so the caption stays through navigation and across recordings. It keeps any newlines in its text.

```{r caption, fig.alt = "The task tracker with a dark caption panel centered along the bottom edge, reading: Seven tasks, two clicks from done.", out.width = "400px"}
page |> pz_annotate_caption("Seven tasks, two clicks from done")
page |> pz_screenshot()
```

paparazzi adds the caption after it crops a still to its frame, so the caption runs along the edge of a framed still, too:

```{r caption-framed, fig.alt = "The task list cropped to its first few tasks, with the dark caption panel along the bottom edge of the crop."}
page |> pz_screenshot(frame = pz_frame(".task-list", pad = 16))
```

paparazzi shows one caption at a time. Each call to `pz_annotate_caption()` replaces the caption on screen, which makes it easy to give each step of a recording its own caption. The caption's reserved id is `"caption"`:

```{r}
page |> pz_annotate_clear(id = "caption")
```

## Clearing and replacing

`pz_annotate_clear()` removes one annotation by `id`, or every annotation when you leave `id` out. Besides the reserved `"spotlight"` and `"caption"`, you can give your own marks, callouts, and redactions any nonempty id. Each call to an annotation function with an id you've used before replaces that annotation instead of adding a second one:

```{r replace, fig.alt = "The new-task form with only a circle around the Add button; the box that was around the title field is gone."}
page |>
  pz_annotate("#task-title", id = "mark") |>
  pz_annotate("#add-task", type = "circle", id = "mark") |>
  pz_screenshot(frame = pz_frame("#new-task", pad = 16))
```

```{r, include = FALSE}
page |> pz_annotate_clear()
```

The second `id = "mark"` call replaced the box around the title field, so only the circle remains.

## Styling defaults

`pz_stage_annotate()` sets your annotation styles once for every later annotation on the page. It sets the color, stroke width, font family, and badge size for every annotation, the bubble colors and distance for callouts, and the side for captions. Arguments to individual annotation functions override these defaults for that call.

Let's make later annotations blue, with large monospace badges:

```{r stage-style, fig.alt = "The Add button boxed in blue with a large monospace A badge."}
page |>
  pz_stage_annotate(
    color = "#2563eb",
    font_family = "monospace",
    font_size = 24
  ) |>
  pz_annotate("#add-task", label = "A") |>
  pz_screenshot(
    frame = pz_frame("#new-task", pad = 8, target_box = "annotated")
  )
```

Staged styles stay on the page until you change them. Passing `NULL` restores a default: crimson (`#e11d48`) marks with 3 pixel lines, sans-serif badges with 14 pixel text, callouts with white text on `#171717`, 24 pixels from their element (8 without a line), and captions at the bottom. We'll clear the box and restore the defaults:

```{r}
page |> pz_annotate_clear()

page |>
  pz_stage_annotate(
    color = NULL,
    font_family = NULL,
    font_size = NULL
  )
```

The staged `fill` and `text_color` style callout bubbles only. Mark badges follow their mark's color unless you pass `label_fill` or `label_text_color`.

### Web fonts

A `font_family` string names a font that Chrome can already find, like a generic family such as `"monospace"` or a font installed on the machine running Chrome. To use a web font, you first load it into the page with `pz_stage_fonts()`. You describe the font with `pz_font_google()`, `pz_font_bunny()`, or, for a font file on disk, `pz_font_file()`. `pz_stage_fonts()` waits until the font has loaded, so paparazzi never draws an annotation in a fallback font while the font downloads.

A handwritten font makes a callout look like a note written over the screenshot. We'll stage Patrick Hand from Google Fonts and make it the default for later annotations:

```{r web-font, fig.alt = "The new-task form with a dark callout below the attachment field, pointing up at it and reading Optional: attach a file in a handwritten font."}
hand <- pz_font_google("Patrick Hand")
page |>
  pz_stage_fonts(hand) |>
  pz_stage_annotate(font_family = hand, font_size = 18) |>
  pz_annotate_callout(
    "Optional: attach a file",
    target = "#attachment",
    side = "bottom"
  ) |>
  pz_screenshot(
    frame = pz_frame("#new-task", pad = 8, target_box = "annotated")
  )
```

You can pass the font object anywhere `font_family` takes a string, as long as you've staged it on the page. It stands for the CSS font family `'"Patrick Hand", sans-serif'`, so the browser falls back to sans-serif for characters the font lacks. Captions and keystroke callouts use the staged `font_family` too, and staged fonts stay with the page when it navigates to a new document.

Chrome downloads Google and Bunny fonts, so staging them needs a network connection. `pz_font_file()` reads the font file in R, so it works offline.

```{r}
page |>
  pz_annotate_clear() |>
  pz_stage_annotate(font_family = NULL, font_size = NULL)
```

## Framing stills

paparazzi measures a still's frame from the elements its target matches. Badges and callouts sit *outside* those elements, so with the default `target_box = "element"`, the frame can crop them off. With `target_box = "annotated"`, paparazzi measures the elements together with their marks, callouts, and spotlight cutouts. Here's the difference on a callout beside the Done filter:

```{r frame-setup}
done <- ".filters a[href='#done']"
page |> pz_annotate_callout("Finished tasks", target = done, side = "right")
```

```{r frame-element, fig.alt = "A tight crop of the Done filter link; the callout arrow beside it is visible but the bubble is cut off."}
page |> pz_screenshot(frame = pz_frame(done, pad = 24))
```

```{r frame-annotated, fig.alt = "A crop of the Done filter link that includes the full callout bubble to its right."}
page |>
  pz_screenshot(
    frame = pz_frame(done, pad = 24, target_box = "annotated")
  )
```

```{r, include = FALSE}
page |> pz_annotate_clear()
```

`target_box = "annotated"` never includes redactions or annotations on other elements. It only applies to stills: paparazzi always measures a recording's frame from the element box.

## Reveals

So far, every annotation has appeared fully drawn. During a recording, paparazzi animates each annotation as it appears, and calls this animation a **reveal**. When you clear an annotation during a recording, paparazzi plays its reveal in reverse. Outside a recording, or while it's paused, paparazzi draws and clears annotations instantly.

`reveal` chooses the animation: `"fade"`, `"draw"` (the outline draws itself), `"pop"`, `"slide"`, `"wipe"` (sweeps clockwise from 12 o'clock), or `"none"`. Boxes fade in by default, the other marks draw themselves, callouts pop in, and spotlights fade in. This recording shows a box fading in, a circle drawing itself, and a callout popping in, then all three reversing out:

```{r reveals}
page |>
  pz_record_start(
    frame = pz_frame("#new-task", pad = 40, when = "start"),
    scale = 0.6,
    hold = c(0.3, 0.5)
  ) |>
  pz_annotate("#task-title", type = "box") |>
  pz_record_hold(0.4) |>
  pz_annotate("#task-priority", type = "circle") |>
  pz_record_hold(0.4) |>
  pz_annotate_callout(
    "Then press Add",
    target = "#add-task",
    side = "bottom"
  ) |>
  pz_record_hold(1) |>
  pz_annotate_clear()
```

The recording chain ends the chunk, so paparazzi stops the recording and adds it to the article as an MP4. In a script, give `pz_record_start()` a path to choose the file.

## Keystroke callouts

`pz_act_press()` can show the keys it presses as keycaps, a callout that paparazzi draws only in recordings. Set [`show_keys`](../reference/pz_act_press.html) to choose how the keycaps look: `"words"` spells out modifier names, `"mac"` uses Mac symbols, and `"both"` shows `Mod` as `Ctrl / ⌘`. The default is `"none"`, and [`pz_stage(show_keys = ...)`](../reference/pz_stage.html) changes it for every later key press on the page.

The task tracker adds a task when you press Enter in the title field, so we'll show Enter with Mac symbols:

```{r show-keys}
page |>
  pz_record_start(
    frame = pz_frame("#new-task", pad = 24, when = "start"),
    scale = 0.6
  ) |>
  pz_act_type("Water the dog", target = "#task-title") |>
  pz_act_press("Enter", show_keys = "mac") |>
  pz_expect_text(
    "Water the dog",
    target = pz_loc(".task-title", which = "first")
  )
```

paparazzi shows one keystroke callout at a time, so each `pz_act_press()` call replaces the previous one. Keystroke callouts never appear in stills.

We're done with the browser, so let's close it:

```{r}
pz_close(page)
```

The [function reference](../reference/index.html#annotations) lists every annotation function with all of its arguments.
