---
title: "Get started with paparazzi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with paparazzi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Suppose you've written a guide to a web app, and you want a short video that shows how to add a task. Doing it by hand means clicking through the app with a screen recorder running, trimming the result, and doing it all again whenever the app changes. With paparazzi, you write the steps down in R instead. paparazzi runs them in a real browser and records the video: the cursor glides to each control, text is typed one character at a time, and the camera zooms in on what matters.

In this article, we'll build this video one step at a time:

```{r destination, echo = FALSE, fig.alt = "A demo of a task tracker. The view zooms in on the new-task form while a cursor types Repot the fern and clicks Add. The view pulls back as the new task appears at the top of the list, outlined in red, under the caption New tasks go to the top of the list."}
library(paparazzi)

demo <- pz_open(
  pz_example("tasks"),
  width = 1000,
  height = 720,
  color_scheme = "light"
)
demo |> pz_stage(enter = "bottom", pause = 0.3)
fern <- pz_loc(".task", has_text = "Repot the fern")

demo |>
  pz_record(
    frame = pz_frame(list("h1", ".task-list"), pad = 24, ratio = 16 / 9),
    format = "gif",
    scale = 800,
    code = {
      demo |>
        pz_annotate_caption("Add a task") |>
        pz_camera("#new-task", wait = TRUE) |>
        pz_act_type("Repot the fern", target = "#task-title") |>
        pz_act_click("#add-task") |>
        pz_camera_reset() |>
        pz_expect_visible(fern) |>
        pz_annotate(fern, type = "box", pad = 4) |>
        pz_annotate_caption("New tasks go to the top of the list") |>
        pz_cursor_leave() |>
        pz_record_hold(1.5)
    }
  )

pz_close(demo)
```

We'll use a small task-tracker page that ships with paparazzi. First we'll open it and take a screenshot. Then we'll script the steps, record them, and add the camera moves, highlight, and captions. At the end, we'll see how the same script works as a test.

You'll need Chrome or another Chromium-based browser, which paparazzi controls through the [chromote](https://rstudio.github.io/chromote/) package. The GIFs in this article need the gifski and png packages, and av for the captions.

```{r setup}
library(paparazzi)
```

## Open a page

`pz_open()` starts a headless browser and loads a page. It takes a URL, a local HTML file, or a Shiny app. `pz_example()` returns the path to the examples that come with paparazzi, so we'll open the task tracker:

```{r}
page <- pz_open(
  pz_example("tasks"),
  width = 1000,
  height = 720,
  color_scheme = "light"
)
page
```

Printing the page shows where the browser is and how it's set up: a 1000 by 720 pixel viewport at twice the usual pixel density, like a laptop's high-resolution screen, with the page's light color scheme. These are **device settings**. They decide what the page looks like, so they also decide what your screenshots and videos look like. You can change them later with `pz_device()`.

`pz_screenshot()` shows us the page:

```{r, fig.alt = "The task tracker, a white card on a gray background, with a form for new tasks, filter links, and a list of tasks."}
page |> pz_screenshot()
```

A screenshot without a file path at the end of a knitted chunk becomes a figure in the document, as above. At the console, it opens in the viewer. Give it a `path` to save a file instead.

## Script the steps

To add a task, a person would type a title into the form and click Add. The script does the same, but it has to say which elements to use. paparazzi finds elements with CSS selectors: the title field is `#task-title` and the Add button is `#add-task`.

**Actions** like `pz_act_type()` and `pz_act_click()` use the page the way a person would, with real mouse and keyboard events. Each action takes the page first and returns it, so a script reads as one chain of steps joined by `|>`:

```{r}
page |>
  pz_act_type("Repot the fern", target = "#task-title") |>
  pz_act_click("#add-task")
```

Before acting, each action waits for its element to be ready: on the page, visible, and not covered by anything else. Then it scrolls the element into view and acts.

Let's check that the new task arrived. `pz_loc()` describes an element more precisely than a selector alone. Here, it's the first `.task-title` on the page, which is where new tasks go. `pz_get_text()` reads that element's text:

```{r}
pz_get_text(page, target = pz_loc(".task-title", which = "first"))
```

That's not what we added. The page shows "Saving…" for a moment before a new task appears, and `pz_get_text()` read the list before the new task arrived. Actions wait for the element they act on, but nothing waited for the result of the click.

For that, we need an **expectation**. Expectation functions start with `pz_expect_`. Each one checks the page and, if the check fails, tries again until it passes or times out:

```{r}
page |>
  pz_expect_text(
    "Repot the fern",
    target = pz_loc(".task-title", which = "first")
  )
```

This time the expectation waited for the page to finish saving. In a demo, expectations do two jobs. They keep the script in step with the page, so the next step doesn't start before the page is ready. And they stop the script if the page doesn't show what you expected, so you never publish a video of the wrong thing. When an expectation times out, its error says what it was looking for and what it last saw:

```{r, error = TRUE}
page |>
  pz_expect_text(
    "Walk the dog",
    target = pz_loc(".task-title", which = "first"),
    timeout = 1
  )
```

A `pz_loc()` description is called a **spec**. A spec doesn't find anything when you create it: paparazzi looks up the elements each time you use it, so one spec keeps working as the page changes. That makes specs good to define once and reuse. We'll describe the new task with one and use it in the video:

```{r}
fern <- pz_loc(".task", has_text = "Repot the fern")
```

`has_text` picks `.task` elements whose text includes "Repot the fern". `pz_loc()` can also pick a match by position with `which`, or look inside another element with `within`.

## Record it

To record a video, wrap the steps in `pz_record()`. It starts recording, runs the code in braces, then stops and writes the file, even if the code throws an error. Leave out the path, as here, and a knitted document includes the recording as a figure, while an interactive session shows it in the viewer. Name the `code` argument when you leave out the path.

We'll reload the page to start fresh, and ask for a GIF, which plays by itself on a web page. `scale = 800` makes it 800 pixels wide:

```{r, fig.alt = "A cursor fades in on the task form, types Repot the fern, and clicks Add. The new task appears at the top of the list."}
page |> pz_nav_reload()

page |>
  pz_record(
    format = "gif",
    scale = 800,
    code = {
      page |>
        pz_act_type("Repot the fern", target = "#task-title") |>
        pz_act_click("#add-task") |>
        pz_expect_visible(fern)
    }
  )
```

The steps are the same ones we ran before, but the recording looks as though a person is using the page. The cursor fades in over the title field and fades out while it types, so it doesn't cover the text. Then it glides to the Add button and presses it. We call this **staging**. It only happens during a recording: outside one, the same steps run at full speed.

`pz_stage()` changes how steps are staged. It can set where the cursor enters, how fast it moves, how big it is, how fast text is typed, and how long to pause after each step. The settings last until you change them. For the final video, the cursor will glide in from the bottom of the frame, and each step will pause for 0.3 seconds so viewers can keep up:

```{r}
page |> pz_stage(enter = "bottom", pause = 0.3)
```

Clicks are staged too. By default, the cursor shrinks briefly as it presses. In a small video that's hard to see, so `pz_act_click()` can draw a ring instead, one that grows out from the click point and fades. Here the cursor checks Urgent with the default press, then unchecks it with `effect = "ring"`:

```{r, fig.alt = "The cursor glides to the Urgent checkbox and checks it with a brief press, then clicks it again, and a crimson ring spreads out from the click as the box unchecks."}
page |>
  pz_record(
    format = "gif",
    scale = 800,
    frame = pz_frame("#new-task", pad = 8),
    code = {
      page |>
        pz_act_click("#task-urgent") |>
        pz_act_click("#task-urgent", effect = "ring")
    }
  )
```

The ring is crimson unless you pass a CSS color as `effect_color`, and `effect = "none"` turns click feedback off. To change every click on the page, set `pz_stage(click_effect = "ring")`, and `click_effect_color` for the ring's color.

## Frame the shot

The recording above shows the whole viewport, with empty space around the card. A **frame** crops a screenshot or recording to part of the page. `pz_frame()` describes one: the elements to fit, the padding around them, and optionally an aspect ratio. We'll fit the card's heading and task list, pad them by 24 pixels, and widen the result to 16:9, the shape of most video players and slides:

```{r}
shot <- pz_frame(list("h1", ".task-list"), pad = 24, ratio = 16 / 9)
```

Like a spec, a frame is only a description. paparazzi measures the page when it uses the frame. A screenshot is a quick way to check one before recording:

```{r, fig.alt = "The task tracker cropped to a 16 by 9 view of the card, from its heading down to the task list."}
page |> pz_screenshot(frame = shot)
```

A frame can also leave out padding or set it separately for each side. See `pz_frame()` for how to anchor the content while it's widened, or keep the frame within another element.

## Direct the viewer's attention

A viewer watching our first recording has to find the form, and then the new task, on their own. The final video shows them both. The **camera** zooms and pans inside the frame, the new task gets an outline, and a **caption** at the bottom of the video says what's happening.

The camera is part of the video, not the page. `pz_camera()` doesn't scroll or zoom the browser, so the steps act on the page exactly as they did before. Outside a recording, it does nothing. Here's the whole script:

```{r, fig.alt = "The view zooms in on the new-task form while a cursor types Repot the fern and clicks Add. The view pulls back as the new task appears at the top of the list, outlined in red, under the caption New tasks go to the top of the list."}
page |> pz_nav_reload()

page |>
  pz_record(
    frame = shot,
    format = "gif",
    scale = 800,
    code = {
      page |>
        pz_annotate_caption("Add a task") |>
        pz_camera("#new-task", wait = TRUE) |>
        pz_act_type("Repot the fern", target = "#task-title") |>
        pz_act_click("#add-task") |>
        pz_camera_reset() |>
        pz_expect_visible(fern) |>
        pz_annotate(fern, type = "box", pad = 4) |>
        pz_annotate_caption("New tasks go to the top of the list") |>
        pz_cursor_leave() |>
        pz_record_hold(1.5)
    }
  )
```

The new lines fall into three groups:

- **Camera.** `pz_camera("#new-task")` zooms in on the form. By default, camera moves run alongside the next step, so the camera and the cursor would move at the same time; `wait = TRUE` lets the camera arrive first. `pz_camera_reset()` pulls back out to the whole frame. While it's zoomed in, the camera also pans to keep each click and keystroke in view.
- **Annotations.** `pz_annotate_caption()` sets the caption, and calling it again replaces the caption. `pz_annotate()` draws a box around the new task, after `pz_expect_visible()` has waited for the task to exist. Annotations follow their elements as the page moves.
- **Pacing.** `pz_cursor_leave()` moves the cursor out of the shot, so it doesn't cover the result. `pz_record_hold()` holds the last frame for 1.5 seconds, so viewers have time to read the caption. The hold is added to the video, not to the script: outside a recording, it doesn't slow anything down.

Annotations stay on the page until you clear them, and they appear in screenshots too. Clear them before moving on:

```{r}
page |> pz_annotate_clear()
```

`pz_annotate()` can also draw circles, underlines, and highlights, with numbered badges. `pz_annotate_callout()` adds text with an arrow, `pz_annotate_spotlight()` dims everything but one element, and `pz_annotate_redact()` covers private information. The [annotations article](annotations.html) shows them all.

## Work in one part of the page

Every task in our list has its own Done button. To mark the fern done, the script needs *its* Done button. `pz_find()` narrows the chain to one element, called the **scope**. After that, targets are looked up inside the scope, so `".task-done"` means this task's Done button:

```{r}
page |>
  pz_find(fern) |>
  pz_act_click(".task-done") |>
  pz_expect_class("done")
```

With no `target`, `pz_expect_class()` checks the scope element itself: the task.

The scope belongs to the chain, not to `page`. `pz_act_click()` changed the browser tab, and every chain on `page` will see the fern marked done. `pz_find()` changed nothing in the browser. It returned a new, scoped context, which went down the chain, while `page` still refers to the whole page. The next chain that starts from `page` starts at the whole page again. To keep working in a scope, assign it to a variable:

```{r}
task_list <- pz_find(page, ".task-list")
task_list
```

Printing a scoped context shows the scope under a `paparazzi scope` header. The Scope line reads from the whole page (`root`) down to the task list, and `(1)` says the scope holds one element. New chains from `task_list` start inside it:

```{r}
pz_get_count(task_list, target = ".task.done")
```

Two tasks are done: the fern, and "Return library books", which starts out done.

## Make other versions

Each version of a demo, like a dark-mode version or a phone version, only needs different device settings. `pz_device()` changes them on an open page. The task tracker follows the browser's color scheme, so here it is on a phone in dark mode:

```{r, fig.alt = "The task tracker on a narrow phone screen, in dark mode.", out.width = "300px"}
page |>
  pz_device(width = 390, height = 700, mobile = TRUE, color_scheme = "dark") |>
  pz_screenshot()
```

`pz_device()` can also zoom the page, turn on reduced motion, or set the locale and time zone. Device settings stay in place until you change them, so a recording made now would also be a phone recording in dark mode.

## When something goes wrong

If an action can't find its element, or finds the wrong one, `pz_inspect()` shows what paparazzi sees. It prints the page summary and the matches for a target, looked up from the current scope. Because it returns the context it was given, you can drop it into the middle of a chain:

```{r}
page |>
  pz_find(".task-list") |>
  pz_inspect(".task-done", show = "none")
```

`.task-done` matches one button for each task. The buttons of finished tasks are hidden, so `pz_inspect()` lists them as `hidden`. An action targeting one of them would wait for it to become visible, then time out. With `show = "screenshot"`, `pz_inspect()` also saves a screenshot with the matches outlined and numbered.

We're done with the browser, so let's close it:

```{r}
pz_close(page)
```

## Browser tests with paparazzi

Staging only happens during a recording, and camera moves and holds do nothing outside one. So if you take away `pz_record()`, the steps of the demo run at full speed, and the expectations still check the page. That's a browser test.

Inside a testthat test, paparazzi's expectations count as test expectations, and a failed expectation is a test failure. `pz_local_page()` opens a page and closes it when the test ends. Here are the steps of our video as a test:

```r
test_that("a new task goes to the top of the list", {
  page <- pz_local_page(pz_example("tasks"))

  page |>
    pz_act_type("Repot the fern", target = "#task-title") |>
    pz_act_click("#add-task") |>
    pz_expect_text(
      "Repot the fern",
      target = pz_loc(".task-title", which = "first")
    )
})
```

## Where to go next

You've now built a demo video from a script: you drove a page with actions, kept the script in step with expectations, and framed, staged, and annotated the recording. From here:

- The [three-part app walkthrough](walkthrough.html) records a series of short MP4 videos with camera moves, a callout, captions, and redacted task titles.
- The [annotations article](annotations.html) shows every kind of annotation, including marks, callouts, spotlights, redactions, and captions, and how to style, frame, and clear them.
- The [Shiny apps article](shiny.html) shows how paparazzi starts Shiny apps, waits for them, and sets their inputs, and how to test them.
- The [function reference](../reference/index.html) lists every function by task, with an example for each.
